The Name of the Title is Hope

### 2023-03-06 input/3_related_work.tex
Updates from Overleaf

\subsection{Disrupted Sense of Agency in Human Augmentation}

This defines the concept of \textbf{natural augmentation} which pairs the idea of wearable designs with reduced or non-existent physical and cognitive burden when using the interface. This notion was illustrated by~\citet{Eden2021-og} which defined natural/true degree of freedom augmentation for movement as a usable additional interface that would be controlled by the user without sacrificing other movements. For example, the control of a robotic arm via a remote control is sacrificing the degree of freedom of our hands. However, a robotic arm that is controlled as if it was an additional identical separate limb would not impair our initial range of potential movements, thus it would be a true augmentation. We could extend this definition saying that any true augmentation would be an enhancement in the way to interact with the environment without impairing any other physical or mental function.

\subsection{Single-trial Classification of the Intent to Interact}

Our idea to overcome the disruption in agency is to use EEG signal reflecting (inter)action intent.

One possible usage scenario of BCIs can be realized by using the Readiness Potential (RP) to de- tect motion intent and trigger actions accordingly. The is a type of potential that has been known for decades (Kornhuber & Deecke, 1965), with research continuously providing evidence of its role in volitional movement preparation (Libet et al., 1993). While it is similar to an event-related potential (ERP), it does not get generated after a stimulus but before the conscious movement starts (Smulders et al., 2012). RP comprises two stages: an early slow stage that begins up to two seconds before the actual movement and a late steep stage that starts about 400 milliseconds before movement. The first stage manifests in the pre-supplementary motor area and transfers to the premotor cortex shortly af- ter. The second stage manifests contralaterally in the primary motor cortex (Shibasaki & Hallett, 2006).
A recent study has shown that the RP is heavily ingrained in the subconscious mechanisms preced- ing movements that people cannot explicitly suppress (Schultze-Kraft et al., 2021). The inability to suppress the RP indicates that it can always be present in the brain waves before motion onset, and it can serve as a reliable predictor of movement.
Having such a predictor for movement can open up new opportunities for the design of BCIs. Shih et al. (2012) defines BCIs as computer-based systems that acquire brain signals, analyze them, and translate them into commands relayed to an output device to carry out the desired action. The purpose of these systems is to detect and quantify the features of brain signals that indicate the user’s intention and translate these features in real-time into device commands that accomplish the user’s intent (Shih et al., 2012). Therefore, using the RP as a feature for prediction motion intent seems a good design for a BCI. Bai et al. (2011) have been the first to successfully set up such an interface and confirmed its feasibility, and Gerrits (2017) has replicated the setup in a follow-up experiment.

\subsection{Levels of Consciousness in HCI}

Along with this idea,~\citet{Jain2020-og} highlighted that numerous types of HCIs for human augmentation remain unexplored. Following their thoughts, they proposed a finer classification of the traditionally defined levels of consciousness (i.e. unconscious, conscious and meta-conscious) by splitting the unconscious into subliminal and pre-conscious and the meta-conscious into meta-cognitive and meta-somatic (Figure \ref{fig:consciousness}). The subliminal level corresponds to what is outside the range of perception of the user and thus a subliminal stimulus do not increase the cognitive load (e.g effect of a vitamin pill intake). A pre-conscious stimulus is below the awareness of the user but it becomes perceivable if the user change its focus toward it (e.g. the side of the road while driving). The conscious level occupies what the user is currently focusing on and thus most of the cognitive load comes from it. The meta-cognitive process is the awareness of the current conscious state (e.g. reflecting on the speed when walking). Finally, the meta-somatic pathway corresponds to the automatic response of the body to external stimulus, thus without intervention of the user (e.g. chills in response to a cold breeze).
This extended classification puts the spot on the new aspects on which a HCI can be designed and how they could use these pathways as new alternatives to contribute to natural human augmentation. Indeed, while numerous HCIs depend on the conscious state requiring the full user's awareness, the achievement of natural enhanced user experience may rely on the exploration of the interaction with other levels of consciousness.

>>>>

We present a bodily integration experience moving the finger of a user after the user intends to do so, but before the user directs its muscles to move.

\subsection{The Disruption in Sense of Agency in Augmented Humans}

\subsection{Assessing the Sense of Agency when Integrating with Computers}
% how can we measure our outcome measure sense of agency?
% - questionnaires
% - intentional binding
% - phenomenological interview

\subsection{Brain-Computer Interfaces for Intent Prediction}
% Single-trial Classification of the Intent to Interact
Our idea to overcome the disruption in agency is to use EEG signals reflecting (inter)action intent. One such signal is the Readiness Potential (RP), or \textit{lateralized readiness potential}. It is an amplitude fluctuation that has been frequently observed preceding voluntary action~\cite{Deecke1969-bl, Libet1983-qu}. It is reliably observed in electrodes placed over the motor cortex contralateral to the acting hand. Since its measurable onset precedes the time of participants self-reported conscious movement intention, it has drawn much interest with respect to the debate on free will, see~\cite{Schurger2021-vp} for a recent neuroscientific perspective. However, evidence abounds for its role in action preparation. An RP is typically comprised of two stages: an early slow stage that begins up to two seconds before the actual movement and a late steep stage that starts about 400 milliseconds before movement. The first stage manifests in the pre-supplementary motor area and transfers to the premotor cortex shortly after. The second stage manifests contra-laterally in the primary motor cortex~\cite{Shibasaki2006-mt}. A recent study has shown that the RP is heavily ingrained in the subconscious mechanisms preceding movements that people cannot explicitly suppress~\cite{Schultze-Kraft2021-cu}. In their study,~\cite{Schultze-Kraft2021-cu} asked participants to find a way to perform a voluntary movement so that the accompanying RP amplitude was as small as possible. After each trial they informed participants about the strength of the RP in the current trial, so participants had a feedback metric to optimize for. They found participants \textit{not} to be able to suppress their RP. This inability to suppress the RP renders it a reliable feature for classification. While its role in movement preparation remains somewhat controversial, it is established as a signal occurring during movement planning and motivates a usage as a reliable predictor of movement.

### 2023-03-06 input/4_user_study_and_methods.tex
Updates from Overleaf

With this study we wanted to find out whether agency can be preserved during \textit{physical actuation} of the user's body when their own brain signals serve as the control signal. Our system leverages the brain signals known to reflect the volitional intent to interact, i.e. the EEG readiness potential, to trigger electrical muscle stimulation to cue a reaching movement.

To assess if agency was preserved, we used a mixed methods approach including a psychometric test of intentional binding, questionnaires as well as qualitative interviews. Our interviews followed a phenomenological structure \todo{add reference from AH review paper mueller}.

@@

\todo{add actual number} Eight participants (\todo{add mean age and sd}) from our local institution participated in our study. All participants were right handed, had normal or corrected to normal vision and had not experienced EMS before. Participants were compensated with 12 Euro per hour of study participation. Prior to their participation, participants were informed of the nature of the experiment, recording and anonymization procedures and signed a consent form approved by a fast track process of the local ethics committee.

@@

\subsubsection{Experiment Design and Procedure}
Participants completed two experimental blocks. The first block consisted of 60 trials and was used to obtain training data for the single-trial classification system outlined below.

In the second block, participants completed 80 trials of the same task and trial order as in the first block. Now, the brain-computer interface (BCI) detecting readiness potentials from real-time EEG data was continuously running in the background. After the target letter was shown to the participants and they were preparing to reach out and press the corresponding key, the BCI was live. Whenever the BCI detected EEG signals to reflect an interaction intent, EMS was activated causing the participants index finger to lift. The second block was then followed by the mixed methods data collection of participants perceived agency.

@@

During the test phase, both classifiers were applied to the real-time data. With the motion data streaming at 90 Hz we subjected the current sample to the motion classifier. For the EEG data streaming at 250 Hz, the EEG data of the last 1000ms was buffered for the selected best channels and 10 windowed means per channel were computed. Identical to the training data, these features were baseline corrected by subtracting the mean from the last 100 ms preceding the last second, i.e. -1100 ms to -1000 ms, resulting in a 200 dimensional baseline corrected spatiotemporal feature vector. Hence, both classifier yielded one output value at each sample point.

The EMS was triggered whenever the EEG classifier predicted the \textit{pre-movement} class with an above .7 certainty. Additionally, EMS was only triggered when the hand was \textit{not} moving. In each trial, as soon as the hand was in motion once after trial start, the EEG-EMS control was deactivated for the rest of the trial. This guaranteed that participants experienced EMS only during movement preparation.

@@

\indent\textbf{(2) Questionnaire.}
In line with~\cite{Hornbaek} we prompted participants with three statements and a 7-point Likert scale centered at 0. The first one "I felt like the device I was using was part of my body" was in relation to body ownership. The following two statements "It felt like I was in control of the movements during the task." and "What is the degree of control you felt?" prompted participants on their experienced agency.

\indent\textbf{(1) Interview.}
We interviewed user's about the experience with the EMS integration. We asked open how questions in a collaborative interview inquiring about the user's lived experience. After prompting user's to recall their experience we started the interview with: "how did it feel like?". If user's made reference themselves to the EMS device we followed up by inquiring: "why do you think it (the EMS trigger) happened?", "What do you think triggered it?". The goal was to learn about the source attribution user's made about the EMS trigger. Since we leveraged the RP, we hypothesized that user's would attribute, at least in part, the EMS trigger to themselves, i.e. their movement intention. Next, we pressed deeper into learning about the cooperative nature between user and system. We asked "Do you believe you did it alone?" and when they made reference to a specific situation where the EMS was triggered we followed up with "What did it feel like, did you had the feeling the system controlled you or did you cooperate?" and "What would have to be there for it to feel like a cooperation?". We closed the interview with two additional exploratory questions: "Did the system influence your performance? If yes, how and why?" as well as asking them about any application scenarios they have in mind.

>>>>

With this study we wanted to find out whether agency can be preserved during \textit{physical actuation} of the user's body when their own brain signals serve as the control signal. To this end, we build a system that leverages the brain signals known to reflect the volitional intent to interact, i.e. the EEG readiness potential, to trigger electrical muscle stimulation to cue a reaching movement.

To assess if agency was preserved, we used a mixed methods approach including a psychometric test of intentional binding, questionnaires as well as qualitative interviews. Our interviews followed a phenomenological structure \todo{add reference from AH review paper mueller}. We t

@@

\todo{add actual number} Ten participants (\todo{add mean age and sd}) from our local institution participated in our study. All participants were right handed, had normal or corrected to normal vision and had not experienced EMS before. Participants were compensated with 12 Euro per hour of study participation. Prior to their participation, participants were informed of the nature of the experiment, recording and anonymization procedures and signed a consent form.

@@

\subsubsection{Experiment Design and Procedure}
Participants completed three experimental blocks. The first block consisted of 60 trials and was used to obtain training data for the single-trial classification system outlined below.

%The condition of the second and third block, see below, were alternated between participants.

In a second block, participants completed 80 trials of the same task and trial order as in the first block. Now, the brain-computer interface (BCI) detecting readiness potentials from real-time EEG data was continuously running in the background. After the target letter was shown to the participants and they were preparing to reach out and press the corresponding key, the BCI was live. Whenever the BCI detected EEG signals to reflect an interaction intent, EMS was activated causing the participants index finger to lift.

Both experimental blocks two and three were followed by a mixed methods data collection of participants' perceived agency.

@@

During the test phase, both classifiers were applied to the real-time data. With the motion data streaming at 90 Hz we subjected the current sample to the motion classifier. For the EEG data streaming at 250 Hz, the EEG data of the last 1000ms was buffered for the selected best channels and 10 windowed means per channel were computed. Identical to the training data, these features were baseline corrected by subtracting the mean from the last 100 ms preceding the last second, i.e. -1100 ms to -1000 ms, resulting in a 200 dimensional baseline corrected spatiotemporal feature vector. Hence, both classifiers yielded one output value at each sample point.

The EMS was triggered whenever the EEG classifier predicted the \textit{pre-movement} class with an above .7 probability. Additionally, EMS was only triggered when the hand was \textit{not} moving. In each trial, as soon as the hand was in motion once after trial start, the EEG-EMS control was deactivated for the rest of the trial. This guaranteed that participants experienced EMS only during movement preparation.

@@

\indent\textbf{(2) Questionnaire.}
In line with~\cite{Hornbaek} we prompted participants with three statements and a 7-point Likert scale centered at 0. The first one ``I felt like the device I was using was part of my body'' was in relation to body ownership. The following two statements ``It felt like I was in control of the movements during the task.'' and ``What is the degree of control you felt?'' prompted participants on their experienced agency.

\indent\textbf{(3) Interview.}
We interviewed user's about the experience with the EMS integration. We asked open how questions in a collaborative interview inquiring about the user's lived experience. After prompting user's to recall their experience we started the interview with: "how did it feel like?". If user's made reference themselves to the EMS device we followed up by inquiring: "why do you think it (the EMS trigger) happened?", "What do you think triggered it?". The goal was to learn about the source attribution user's made about the EMS trigger. Since we leveraged the RP, we hypothesized that user's would attribute, at least in part, the EMS trigger to themselves, i.e. their movement intention. Next, we pressed deeper into learning about the cooperative nature between user and system. We asked "Do you believe you did it alone?" and when they made reference to a specific situation where the EMS was triggered we followed up with "What did it feel like, did you had the feeling the system controlled you or did you cooperate?" and "What would have to be there for it to feel like a cooperation?". We closed the interview with two additional exploratory questions: "Did the system influence your performance? If yes, how and why?" as well as asking them about any application scenarios they have in mind.
